---
output: github_document
---
<!-- sector-update.md is generated from sector-update.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(dplyr)
library(implan)
```

## Overview

The implan package includes some data and functions to help update category-sector crosswalks between implan sectoring schemes (e.g., [version 536 to 546](https://implanhelp.zendesk.com/hc/en-us/articles/360034896614-546-Industries-Conversions-Bridges-Construction-2018-Data)). Note that the included function provides a direct crosswalk update and doesn't do any respecification (for example if the market basket of goods changes).

#### Master Sector Schemes

The master sectoring schemes (category to sector) are stored on Office 365: [Implan > Sectoring Schemes](https://southwickassociatesinc.sharepoint.com/:f:/s/Implan/EnL9nwfijuROuo34BCV47fkB9yuOh8l0lDdaEtmEfL9TNA?e=Pkiaai).

## Pull Data

The 536to546 crosswalk file has been included in this package. I did some pre-processing to get the crosswalk data into a more useful format ([code here](https://github.com/southwick-associates/implan/blob/master/data-raw/sector536_to_sector546.R)).

```{r}
library(dplyr)
library(implan)

data("sector536_to_sector546")
glimpse(sector536_to_sector546)
```

Note that the `CewAvgRatio` variable from the raw data was renamed to `crosswalk_ratio`. This variable was used for the small number of cases when 536 sectors were expanded to multiple sectors in the 546 scheme.

```{r}
check_share_sums(sector536_to_sector546, crosswalk_ratio, group, sector536)
```

We need the 536 sectoring scheme for the update, an example of which has been included in this package:

```{r}
data("category_to_sector536")
glimpse(category_to_sector536)
```

The `sector_update()` function converts from one sectoring scheme to the next. 

```{r}
data("sectors546")
category_to_sector546 <- sector_update(
    scheme_old = category_to_sector536, 
    crosswalk = sector536_to_sector546, 
    description_new = sectors546
)
glimpse(category_to_sector546)
```

